Pulling oral and IV data from CvTdb

### 2025-08-01 data-raw/pulling_oral_iv_data_cvtdb.Rmd

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      comment = NA,
                      tidy.opts = list(width.cutoff = 75),
                      tidy = TRUE)
```

@@

```{r query_functions}
postgreSQL_connection <- function() {
  return(
    dbConnect(RPostgreSQL::PostgreSQL(),
              user = Sys.getenv("CVT_USER"),
              password = Sys.getenv("CVT_PASS"),
              host = Sys.getenv("CVT_HOST"),
              dbname = Sys.getenv("CVT_DB_NAME"))
  )
}

query_db <- function(query = NULL, schema) {
  if (is.null(query)) return(cat("\nMust provide a query to send"))
  con <- postgreSQL_connection()

  query_result <- tryCatch(
    return(dbGetQuery(con, query %>%
                        gsub("FROM ", paste0("FROM ", schema, "."), .) %>%
                        gsub("JOIN (?=\\w)", paste0("JOIN ", schema, "."), ., perl = TRUE))),
  error = function(cond){ 
    cat(paste0("\nError message: ", cond))
    return(NULL)
  },
  finally = { 
    dbDisconnect(con)
    }
    )
  return(query_result)
}

@@

  for (this_text in seq_along(text)) {
    cli::cli_li(text[this_text])
  }
  cli::cli_end()

}

quick_stats <- function(.data) {
  stopifnot(is.data.frame(.data))
  cli::cli_h2("Statistics on the current data")
  message(paste0("Total number of data points: ", nrow(.data)))
  message("Unique Species and data counts:")
  print(table(.data$species))

  message(paste0("Number of unique studies: ",
                 length(unique(.data$fk_study_id))))
  message(paste0("Number of unique series: ",
                 length(unique(.data$fk_series_id))))
  message(paste0("Number of unique chemicals: ",
                 length(unique(.data$analyte_dtxsid))))
  message(paste0("Unique Dose units: ",
                 toString(unique(.data$dose_level_units_original))))
  message(paste0("Unique Concentration units: ", 
                 toString(unique(.data$conc_units_original))))
  message("Data by unique route of administration:")
  print(table(.data$administration_route_normalized))
  message("Data by medium collected:")
  print(table(.data$conc_medium_normalized))
  message("Data by curation set:")
  print(table(.data$curation_set_tag))
}

@@

```{r diff_log_function}
# Now I want to write a quick function that will log relevant changes
# from previous 'cvt' objects
devtools::load_all() # So this loads the previous cvt object
diff_log_cvt <- function(new_cvt,
                         key_id = "conc_time_id",
                         grp_ids = c("analyte_dtxsid",
                                     "species"),
                         notes = character(0L)) {
  stopifnot(
    key_id %in% colnames(cvt),
    key_id %in% colnames(new_cvt),
    all(grp_ids %in% colnames(cvt)),
    all(grp_ids %in% colnames(new_cvt))
  )
  if (length(unique(new_cvt[[key_id]])) != NROW(new_cvt)) {
    stop("key_id is not not a column that uniquely identifies this data.")
  }

@@

  if (identical(new_cvt, cvt)) {
    msg_to_write <- paste(msg_to_write, "No changes.", sep = "\t")
    message("There are no changes from previous `cvt` data.frame, ",
            "are you sure you want to save this?")
  } else {
    msg_to_write <- paste(msg_to_write, "Changes detected.", sep = "\t")

    ###---- Check key_id changes (usually conc_time_id)
    diff_row_add <- setdiff(new_cvt[[key_id]], cvt[[key_id]])
    diff_row_sub <- setdiff(cvt[[key_id]], new_cvt[[key_id]])
    diff_row_common <- intersect(cvt[[key_id]], new_cvt[[key_id]])
    # Nature of data changes:
    added_grps <- unique(
      do.call("paste",
              new_cvt[which(new_cvt[[key_id]] %in% diff_row_add), grp_ids]
              )
      )
    removed_grps <- unique(
      do.call("paste",
              cvt[which(cvt[[key_id]] %in% diff_row_sub), grp_ids]
              )
      )

@@

    limit_grp_output <- function(grp_vector) {
      if (length(grp_vector) > 0) {
        if (length(grp_vector) > 6) {
          grp_vector <- c(grp_vector[c(1,2)],
                              paste0("... [",
                                     length(grp_vector) - 3,
                                     "] more!"),
                              grp_vector[c(length(grp_vector) - 1)]
          )
        }
      return(grp_vector)
    }
      msg_to_write <- c(msg_to_write,
                        "New Groups added (not in previous data):",
                        paste0("\t", limit_grp_output(added_new_grps)),
                        "Groups with new observations (group present in previous data)",
                        paste0("\t", limit_grp_output(added_existing_grps)),
                        "Groups removed that no longer have data present:",
                        paste0("\t", limit_grp_output(removed_new_grps)),
                        "Groups removed that still have some data present:",
                        paste0("\t", limit_grp_output(removed_existing_grps)),
                        "-----------------------------------"
                        )
    }

@@

    if (length(c(diff_col_add, diff_col_sub)) == 0) {
      msg_to_write <- c(msg_to_write,
                        "No changes in column names.")
    } else if (length(diff_col_add) > 0 && length(diff_col_sub) > 0) {
      # Check renaming

      diff_columns <- function(x, y) {
        out_list <- character()
        for (ix in setdiff(names(x), key_id)) {
          for (iy in setdiff(names(y), key_id)) {
            tmp <- merge(x[c(key_id,ix)], y[c(key_id, iy)], by=key_id)
            if (identical(tmp[[2]], tmp[[3]])) {
              out_list <- c(out_list, paste0(ix, " -> ", iy))
            }
          }
        }
        return(out_list)
      }

      renamed_cols <- diff_columns(common_row_cvt[c(key_id,diff_col_sub)],
                                   common_row_new[c(key_id, diff_col_add)])
      if (length(renamed_cols) > 0) {
        msg_to_write <- c(msg_to_write,
                          "Renamed columns (non-exhaustive):",
                          renamed_cols)
      } else {
        msg_to_write <- c(msg_to_write,
                          "No renamed columns found.")
      }
    }
    if (length(diff_col_add) > 0) {
      msg_to_write <- c(msg_to_write,
                        "--------------",
                        "Columns added:",
                        diff_col_add)
    }
    if (length(diff_col_sub) > 0) {
      msg_to_write <- c(msg_to_write,
                        "______________",
                        "Columns removed (from previous data):",
                        diff_col_sub)
    }

    if ((NROW(
      setdiff(
        common_row_cvt[diff_col_common], common_row_new[diff_col_common]
      )) > 0
    ) ||  (NROW(
      setdiff(
        common_row_new[diff_col_common], common_row_cvt[diff_col_common]
      )) > 0
    )) {
      msg_to_write <- c(msg_to_write,
                        "Data for same observation key_ids have changed.")

      # Which columns have different data?
      test_cols <- setdiff(diff_col_common, key_id)
      for (i in test_cols) {
        this_diff <- setdiff(common_row_new[c(key_id, i)],
                             common_row_cvt[c(key_id, i)])
        that_diff <- setdiff(common_row_cvt[c(key_id, i)],
                             common_row_new[c(key_id, i)])

@@

          this_grp <- dplyr::distinct(common_row_new[
            common_row_new[[key_id]] %in% this_diff[[key_id]],
            grp_ids
          ])
          this_grp <- paste0(this_grp[1,1:2], collapse = " ")

          # Print the differences (if below a certain number of rows)
          msg_to_write <- c(
            msg_to_write,
            paste0(
              "Example:",
              paste(this_grp, collapse = " "),
              collapse = " "
            ),
            paste0(c(i, "\t new:", this_diff[1,2], "<--\t old:", that_diff[1,2]),
                   collapse = " "), "\n"
          )
        }
      }
    }
    msg_to_write <- c(msg_to_write,
                      notes,
                      "================================\n")
  }
  return(cat(msg_to_write, sep = "\n"))
}

@@

## Suggestions for CvTdb

Repairing duplication in the chemicals table should be a priority
that would benefit all current and future users of _CvTdb_. It would repair other
missing identifying chemical information as well as truly have unique rows per chemical.
Here this repair is done by making names lowercase to avoid capitalization, but there
may be other issues arise with spaces that should be parsed prior to input into 
the database record.

The other suggestion would be to have a filled out `conc_units_normalized` column in
the `series` table. Although it is assumed to be _mg/kg_, it would be nice to have this
confirmed. Here `conc_units_original` is compared with the ratio of `conc` to `conc_original`
to get a sense of whether the units for `conc` were reasonably converted to
_mg/kg_. This seems like it could reasonably be a part of the QA process.

Besides these two suggestions, there are only a couple other changes in the data that 
are made throughout this document but these pertain to specific records rather than
the entire database, so I think they are a little bit lower on the priority list.

```{r SQL_queries, warning=FALSE}
# Need to make this in a specified ordered
# First all IDs (so I can check them quickly)
# Note that this uses an ALIAS for each table which is later specified in
# the FROM and LEFT JOIN statments
query <- paste0(
  "SELECT distinct ",
  # Conc_Time_Values table fields
  "a.id, a.fk_series_id, a.time_original, a.time_hr, ",
  "a.conc_original, a.conc_sd_original, a.conc, a.conc_sd, ",

  # Series table fields
  "b.fk_analyzed_chemical_id, ",
  ## Chemical dictionary fields (analyzed chemical information)
  "l.analyzed_chem_dtxsid, l.analyzed_chem_name_original, ",
  "l.analyzed_chem_casrn, l.analyzed_chem_name, ",

  "b.time_units_original, b.conc_units_original, ",
  "b.conc_units_normalized, b.loq, b.loq_units, ",
  "b.n_subjects_in_series, b.radiolabeled, b.fk_study_id, ",

  # Studies table fields
  "c.fk_dosed_chemical_id, ",
  ## Chemical dictionary fields (dosed chemical information)
  "k.dosed_chem_dtxsid, k.dosed_chem_name_original, ",
  "k.dosed_chem_casrn, k.dosed_chem_name, ",

  "c.dose_volume, c.dose_volume_units, c.dose_vehicle, ",
  "c.dose_duration, c.dose_duration_units, ",
  ## dose_frequency dictionary fields
  "j.dose_frequency_original, j.dose_frequency_normalized, ",
  "c.fasting_period, ",
  "c.dose_level_normalized, c.dose_level_original, c.dose_level_units_original, ",
  ## conc_medium dictionary fields
  "i.conc_medium_original, i.conc_medium_normalized, ", 
  ## administration_route dictionary fields
  "h.administration_route_original, h.administration_route_normalized, ",
  ## administration_method dictionary fields
  "g.administration_method_original, g.administration_method_normalized, ",
  ## administration_form dictionary fields
  "f.administration_form_original, f.administration_form_normalized, ",

@@

  # Join to dictionary tables
  "LEFT JOIN administration_form_dict f ON c.fk_administration_form_id = f.id ",
  "LEFT JOIN administration_method_dict g ON c.fk_administration_method_id = g.id ",
  "LEFT JOIN administration_route_dict h ON c.fk_administration_route_id = h.id ",
  "LEFT JOIN conc_medium_dict i ON b.fk_conc_medium_id = i.id ",
  "LEFT JOIN dose_frequency_dict j ON c.fk_dose_frequency_id = j.id ",

  # Rename chemical fields for dosed vs. analyzed chemical record foreign keys
  "LEFT JOIN (SELECT id, dsstox_substance_id as dosed_chem_dtxsid, ",
  "chemical_name_original as dosed_chem_name_original, ", 
  "dsstox_casrn as dosed_chem_casrn, preferred_name as dosed_chem_name ",
  "FROM chemicals) as k ON c.fk_dosed_chemical_id = k.id ",
  "LEFT JOIN (SELECT id, dsstox_substance_id as analyzed_chem_dtxsid, ",
  "chemical_name_original as analyzed_chem_name_original, ",
  "dsstox_casrn as analyzed_chem_casrn, preferred_name as analyzed_chem_name ",
  "FROM chemicals) as l ON b.fk_analyzed_chemical_id = l.id ",

  # Filtering steps
  "WHERE h.administration_route_normalized IN ('iv', 'oral') AND ",
  "(i.conc_medium_normalized IN ('blood', 'plasma', 'excreta') OR ",
  "i.conc_medium_original IN ('blood', 'plasma', 'excreta'))"
)

@@

# Need to change column names for this
{
  names(sql_query)[1] <- "conc_time_id"
}

# Unfortunately, there are some "fancy" or curly quotes that need to be replaced
# with "straight" quotes to be more machine read-able
force_straight_quotes <- function(x) {
  temp <- gsub("(’|′)", "'", x)
  temp <- gsub("(, | ,)", ",", temp)
  return(temp)
}
force_straight_quotes("3,3’,4,4’-tetrachloro-5-biphenylol")

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  comment = NA,
  tidy.opts = list(width.cutoff = 75),
  tidy = TRUE
)
```

@@

```{r query_functions}
postgreSQL_connection <- function() {
  return(
    dbConnect(RPostgreSQL::PostgreSQL(),
      user = Sys.getenv("CVT_USER"),
      password = Sys.getenv("CVT_PASS"),
      host = Sys.getenv("CVT_HOST"),
      dbname = Sys.getenv("CVT_DB_NAME")
    )
  )
}

query_db <- function(query = NULL, schema) {
  if (is.null(query)) {
    return(cat("\nMust provide a query to send"))
  }
  con <- postgreSQL_connection()

  query_result <- tryCatch(
    return(dbGetQuery(con, query %>%
      gsub("FROM ", paste0("FROM ", schema, "."), .) %>%
      gsub("JOIN (?=\\w)", paste0("JOIN ", schema, "."), .,
        perl = TRUE
      ))),
    error = function(cond) {
      cat(paste0("\nError message: ", cond))
      return(NULL)
    },
    finally = {
      dbDisconnect(con)
    }
  )
  return(query_result)
}
```

@@

  for (this_text in seq_along(text)) {
    cli::cli_li(text[this_text])
  }
  cli::cli_end()
}

quick_stats <- function(.data, old_version = FALSE) {
  stopifnot(is.data.frame(.data))

  cli::cli_h2("Statistics on the current data")
  message(paste0("Total number of data points: ", nrow(.data)))
  message("Unique Species and data counts:")
  print(table(.data$species))

  message(
    "Number of unique studies: ",
    length(unique(.data$fk_study_id))
  )
  message(
    "Number of unique series: ",
    length(unique(.data$fk_series_id))
  )

  if (old_version) {
    message(
      "Number of unique chemicals: ",
      length(unique(.data$analyte_dtxsid)),
      "\n"
    )
    chem_by_medium <- .data[c("conc_medium_normalized", "analyte_dtxsid")] %>%
      filter(!is.na(analyte_dtxsid)) %>%
      distinct()
    chem_by_curset <- .data[c("curation_set_tag", "analyte_dtxsid")] %>%
      filter(!is.na(analyte_dtxsid)) %>%
      distinct()
  } else {
    message(
      "Number of unique chemicals: ",
      length(unique(.data$analyzed_chem_dtxsid)),
      "\n"
    )
    chem_by_medium <- .data[c("conc_medium_normalized", "analyzed_chem_dtxsid")] %>%
      filter(!is.na(analyzed_chem_dtxsid)) %>%
      distinct()
    chem_by_curset <- .data[c("curation_set_tag", "analyzed_chem_dtxsid")] %>%
      filter(!is.na(analyzed_chem_dtxsid)) %>%
      distinct()
  }

  message(
    "Unique Dose units: ",
    toString(unique(.data$dose_level_units_original)),
    "\n"
  )
  if (all(is.na(.data$conc_units_normalized))) {
    message(
      "Unique Concentration units: ",
      toString(unique(.data$conc_units_original)),
      "\n"
    )
  } else {
    message(
      "Unique Concentration units: ",
      toString(unique(.data$conc_units_normalized)),
      "\n"
    )
  }
  message("Data by unique route of administration:")
  print(table(.data$administration_route_normalized))
  message("Data by medium collected:")
  print(table(.data$conc_medium_normalized))
  message("Data by curation set:")
  print(table(.data$curation_set_tag))
  message("\n\nUnique chemicals per medium:")
  print(table(chem_by_medium$conc_medium_normalized))
  message("Unique chemicals per curation set:")
  print(table(chem_by_curset$curation_set_tag))
}

chem_id_cols <- rlang::quos(
  fk_analyzed_chemical_id, fk_dosed_chemical_id,
  analyzed_chem_dtxsid, dosed_chem_dtxsid,
  analyzed_chem_name, dosed_chem_name,
  analyzed_chem_name_original, dosed_chem_name_original,
  analyzed_chem_name_secondary_original, 
  dosed_chem_name_secondary_original
)

unit_values_cols <- rlang::quos(
  conc_original, conc,
  conc_units_original, conc_units_normalized,
  conc_sd_original, conc_sd,
  loq, loq_units,
  dose_level_original, dose_level_normalized,
  dose_level_units_original, dose_volume_units,
  dose_volume, weight_kg
)

@@

```{r diff_log_function}
# Now I want to write a quick function that will log relevant changes
# from previous 'cvt' objects
diff_log_cvt <- function(new_cvt,
                         key_id = "conc_time_id",
                         grp_ids = c(
                           "analyzed_chem_dtxsid",
                           "species"
                         ),
                         notes = character(0L)) {
  stopifnot(
    key_id %in% colnames(cvt),
    key_id %in% colnames(new_cvt),
    all(grp_ids %in% colnames(cvt)),
    all(grp_ids %in% colnames(new_cvt))
  )
  if (length(unique(new_cvt[[key_id]])) != NROW(new_cvt)) {
    stop("key_id is not not a column that uniquely identifies this data.")
  }

@@

  if (identical(new_cvt, cvt)) {
    msg_to_write <- paste(msg_to_write, "No changes.", sep = "\t")
    message(
      "There are no changes from previous `cvt` data.frame, ",
      "are you sure you want to save this?"
    )
  } else {
    msg_to_write <- paste(msg_to_write, "Changes detected.", sep = "\t")

    ### ---- Check key_id changes (usually conc_time_id)
    diff_row_add <- setdiff(new_cvt[[key_id]], cvt[[key_id]])
    diff_row_sub <- setdiff(cvt[[key_id]], new_cvt[[key_id]])
    diff_row_common <- intersect(cvt[[key_id]], new_cvt[[key_id]])
    # Nature of data changes:
    added_grps <- unique(
      do.call(
        "paste",
        new_cvt[which(new_cvt[[key_id]] %in% diff_row_add), grp_ids]
      )
    )
    removed_grps <- unique(
      do.call(
        "paste",
        cvt[which(cvt[[key_id]] %in% diff_row_sub), grp_ids]
      )
    )

@@

    limit_grp_output <- function(grp_vector) {
      if (length(grp_vector) > 0) {
        if (length(grp_vector) > 6) {
          grp_vector <- c(
            grp_vector[c(1, 2)],
            paste0(
              "... [",
              length(grp_vector) - 3,
              "] more!"
            ),
            grp_vector[c(length(grp_vector) - 1)]
          )
        }
        return(grp_vector)
      }
      msg_to_write <- c(
        msg_to_write,
        "New Groups added (not in previous data):",
        paste0("\t", limit_grp_output(added_new_grps)),
        "Groups with new observations (group present in previous data)",
        paste0("\t", limit_grp_output(added_existing_grps)),
        "Groups removed that no longer have data present:",
        paste0("\t", limit_grp_output(removed_new_grps)),
        "Groups removed that still have some data present:",
        paste0("\t", limit_grp_output(removed_existing_grps)),
        "-----------------------------------"
      )
    }

@@

    if (length(c(diff_col_add, diff_col_sub)) == 0) {
      msg_to_write <- c(
        msg_to_write,
        "No changes in column names."
      )
    } else if (length(diff_col_add) > 0 && length(diff_col_sub) > 0) {
      # Check renaming

      diff_columns <- function(x, y) {
        out_list <- character()
        for (ix in setdiff(names(x), key_id)) {
          for (iy in setdiff(names(y), key_id)) {
            tmp <- merge(x[c(key_id, ix)], y[c(key_id, iy)], by = key_id)
            if (identical(tmp[[2]], tmp[[3]])) {
              out_list <- c(out_list, paste0(ix, " -> ", iy))
            }
          }
        }
        return(out_list)
      }

      renamed_cols <- diff_columns(
        common_row_cvt[c(key_id, diff_col_sub)],
        common_row_new[c(key_id, diff_col_add)]
      )
      if (length(renamed_cols) > 0) {
        msg_to_write <- c(
          msg_to_write,
          "Renamed columns (non-exhaustive):",
          renamed_cols
        )
      } else {
        msg_to_write <- c(
          msg_to_write,
          "No renamed columns found."
        )
      }
    }
    if (length(diff_col_add) > 0) {
      msg_to_write <- c(
        msg_to_write,
        "--------------",
        "Columns added:",
        diff_col_add
      )
    }
    if (length(diff_col_sub) > 0) {
      msg_to_write <- c(
        msg_to_write,
        "______________",
        "Columns removed (from previous data):",
        diff_col_sub
      )
    }

    if ((NROW(
      setdiff(
        common_row_cvt[diff_col_common], common_row_new[diff_col_common]
      )
    ) > 0
    ) || (NROW(
      setdiff(
        common_row_new[diff_col_common], common_row_cvt[diff_col_common]
      )
    ) > 0
    )) {
      msg_to_write <- c(
        msg_to_write,
        "Data for same observation key_ids have changed."
      )

      # Which columns have different data?
      test_cols <- setdiff(diff_col_common, key_id)
      for (i in test_cols) {
        this_diff <- setdiff(
          common_row_new[c(key_id, i)],
          common_row_cvt[c(key_id, i)]
        )
        that_diff <- setdiff(
          common_row_cvt[c(key_id, i)],
          common_row_new[c(key_id, i)]
        )

@@

          this_grp <- dplyr::distinct(common_row_new[
            common_row_new[[key_id]] %in% this_diff[[key_id]],
            grp_ids
          ])
          this_grp <- paste0(this_grp[1, 1:2], collapse = " ")

          # Print the differences (if below a certain number of rows)
          msg_to_write <- c(
            msg_to_write,
            paste0(
              "Example:",
              paste(this_grp, collapse = " "),
              collapse = " "
            ),
            paste0(c(i, "\t new:", this_diff[1, 2], "<--\t old:", that_diff[1, 2]),
              collapse = " "
            ), "\n"
          )
        }
      }
    }
    msg_to_write <- c(
      msg_to_write,
      notes,
      "================================\n"
    )
  }
  return(cat(msg_to_write, sep = "\n"))
}
```

@@

```{r SQL_queries, warning=FALSE}
# Need to make this in a specified ordered
# First all IDs (so I can check them quickly)
# Note that this uses an ALIAS for each table which is later specified in
# the FROM and LEFT JOIN statements
series_chems <- paste0(
  "SELECT DISTINCT a.id, a.fk_analyzed_chemical_id, ",
  "b.analyzed_chem_dtxsid, b.analyzed_chem_name_original, ",
  "b.analyzed_chem_name_secondary_original, ",
  "b.analyzed_chem_casrn, b.analyzed_chem_name, ",
  "a.time_units_original, a.conc_units_original, ",
  "a.conc_units_normalized, a.loq, a.loq_units, ",
  "a.n_subjects_in_series, a.radiolabeled, a.fk_study_id, ",
  "a.qc_notes, a.qc_status, a.qc_flags, c.conc_medium_normalized ",
  "FROM series a ",
  "LEFT JOIN (SELECT id, dsstox_substance_id AS analyzed_chem_dtxsid, ",
  "chemical_name_original AS analyzed_chem_name_original, ",
  "chemical_name_secondary_original AS analyzed_chem_name_secondary_original, ",
  "dsstox_casrn as analyzed_chem_casrn, preferred_name AS analyzed_chem_name ",
  "FROM chemicals) AS b ON a.fk_analyzed_chemical_id = b.id ",
  "LEFT JOIN conc_medium_dict AS c ON a.fk_conc_medium_id = c.id"
)

study_docs <- paste0(
  "SELECT a.id, a.fk_extraction_document_id, a.fk_dosed_chemical_id, ",
  "b.pmid, b.doi, b.year, b.other_study_identifier, ",
  "b.url, b.extracted, b.curation_set_tag ",
  "FROM studies a ",
  "LEFT JOIN documents b ON a.fk_extraction_document_id = b.id"
)

audit_query <- query_db(
  paste0(
    "SELECT DISTINCT * FROM cvt_audit ",
    "WHERE fk_table_name IN ('documents', 'series', 'studies')"
  ),
  "cvt"
)
study_query <- query_db(study_docs, "cvt")
series_query <- query_db(series_chems, "cvt")
dtxsid_query <- query_db("SELECT DISTINCT * FROM chemicals", "cvt") %>%
  mutate(
    across(ends_with("_id"), as.character),
    across(contains("_name"), \(x) {
      stringr::str_to_lower(x) |>
        force_ascii_quotes() |>
        # trim whitespace around hyphens
        stringr::str_replace_all("\\s*-\\s*", "-") |>
        # comma-separate forms like 3'6 into 3',6
        stringr::str_replace_all("(?<=\\d)'(?=\\d)", "',") |>
        # fix some bracket cases
        stringr::str_replace_all("\\[3\\]h", "[3h]") |>
        stringr::str_replace_all("benzo\\[a\\]", "benzo(a)") |>
        # adds a space after square bracket separator
        stringr::str_replace_all("](?!\\s|-)", "] ")
    })
  )

select_chems <- subset(
  dtxsid_query,
  select = c(
    "id",
    "dsstox_substance_id",
    "dsstox_casrn",
    "preferred_name",
    "chemical_name_original",
    "chemical_name_secondary_original"
  )
)

# Surveying the non-equal chemical id landscape:
series_query |>
  inner_join(study_query, by = join_by(fk_study_id == id)) |>
  distinct(fk_analyzed_chemical_id, fk_dosed_chemical_id, curation_set_tag) |>
  count(fk_analyzed_chemical_id == fk_dosed_chemical_id)

unequal_chem_ids <- series_query |>
  inner_join(study_query, by = join_by(fk_study_id == id)) |>
  distinct(fk_analyzed_chemical_id, fk_dosed_chemical_id, curation_set_tag) |>
  filter(fk_analyzed_chemical_id != fk_dosed_chemical_id) |>
  left_join(
    rename_with(select_chems, ~ paste0("analyzed_", .x)),
    by = join_by(fk_analyzed_chemical_id == analyzed_id)
  ) |>
  left_join(
    rename_with(select_chems, ~ paste0("dosed_", .x)),
    by = join_by(fk_dosed_chemical_id == dosed_id)
  ) |>
  glimpse()

count(
  unequal_chem_ids,
  analyzed_dsstox_substance_id == dosed_dsstox_substance_id,
  analyzed_dsstox_casrn == dosed_dsstox_casrn,
  analyzed_preferred_name == dosed_preferred_name,
  analyzed_chemical_name_original == dosed_chemical_name_original,
  analyzed_chemical_name_secondary_original == dosed_chemical_name_secondary_original
) |>
  rename(
    "Equal DTXSID" = 1, "Equal CASRN" = 2, "Equal Preffered Name" = 3,
    "Equal Recorded Name" = 4, "Equal Secondary Recorded Name" = 5, "N" = 6
  )

# Are any names equal???
count(
  unequal_chem_ids,
  analyzed_dsstox_substance_id == dosed_dsstox_substance_id |
    analyzed_dsstox_casrn == dosed_dsstox_casrn |
    analyzed_preferred_name == dosed_preferred_name |
    analyzed_chemical_name_original == dosed_chemical_name_original |
    analyzed_chemical_name_secondary_original == dosed_chemical_name_secondary_original
) |>
  rename(
    "Any" = 1, "N" = 2
  ) |>
  mutate(
    Any = ifelse(Any %in% TRUE, "Some Identical", "None Identical")
  ) |>
  group_by(Any) |>
  summarize(N = sum(N)) |>
  pivot_wider(names_from = Any, values_from = N)

# Let's parse through with those "Some Identical" identifiers first
unequal_chem_ids |>
  filter(
    analyzed_dsstox_substance_id == dosed_dsstox_substance_id |
      analyzed_dsstox_casrn == dosed_dsstox_casrn |
      analyzed_preferred_name == dosed_preferred_name |
      analyzed_chemical_name_original == dosed_chemical_name_original |
      analyzed_chemical_name_secondary_original == dosed_chemical_name_secondary_original
  ) |>
  group_by(fk_analyzed_chemical_id, fk_dosed_chemical_id) |>
  summarize(
    same_dtxsid = if_else(
      analyzed_dsstox_substance_id == dosed_dsstox_substance_id,
      analyzed_dsstox_substance_id, NA_character_, missing = NA_character_
    ),
    same_casrn = if_else(
      analyzed_dsstox_casrn == dosed_dsstox_casrn,
      analyzed_dsstox_casrn, NA_character_, NA_character_
    ),
    same_preferred_name = if_else(
      analyzed_preferred_name == dosed_preferred_name,
      analyzed_preferred_name, NA_character_, NA_character_
    ),
    same_chem_name = if_else(
      analyzed_chemical_name_original == dosed_chemical_name_original,
      analyzed_chemical_name_original, NA_character_, NA_character_
    ),
    same_chem_name_secondary = if_else(
      analyzed_chemical_name_secondary_original == dosed_chemical_name_secondary_original,
      analyzed_chemical_name_secondary_original, NA_character_, NA_character_
    )
  ) |>
  rowwise() |>
  mutate(
    likely_same = sum(!is.na(c_across(starts_with("same_"))))
  ) |>
  ungroup() |>
  arrange(same_dtxsid, same_casrn, same_preferred_name, same_chem_name,
          same_chem_name_secondary) |>
  View("Similar IDs")
# There are two ways the identifiers may be duplicated,
# shown here between rows and columns

# There are a lot of duplicate DTXSIDs
query <- paste0(
  "SELECT distinct ",
  # Conc_Time_Values table fields
  "a.id, a.fk_series_id, a.time_original, a.time_hr, ",
  "a.conc_original, a.conc_sd_original, a.conc, a.conc_sd, ",

  # Series table fields
  "b.fk_analyzed_chemical_id, ",
  ## Chemical dictionary fields (analyzed chemical information)
  "l.analyzed_chem_dtxsid, l.analyzed_chem_name_original, ",
  "l.analyzed_chem_casrn, l.analyzed_chem_name, ",
  "l.analyzed_chem_name_secondary_original, ",
  "b.time_units_original, b.conc_units_original, ",
  "b.conc_units_normalized, b.loq, b.loq_units, ",
  "b.n_subjects_in_series, b.radiolabeled, b.fk_study_id, ",

  # Studies table fields
  "c.fk_dosed_chemical_id, ",
  ## Chemical dictionary fields (dosed chemical information)
  "k.dosed_chem_dtxsid, k.dosed_chem_name_original, ",
  "k.dosed_chem_casrn, k.dosed_chem_name, ",
  "k.dosed_chem_name_secondary_original, ",
  "c.dose_volume, c.dose_volume_units, c.dose_vehicle, ",
  "c.dose_duration, c.dose_duration_units, ",
  ## dose_frequency dictionary fields
  "j.dose_frequency_original, j.dose_frequency_normalized, ",
  "c.fasting_period, ",
  "c.dose_level_normalized, c.dose_level_original, ",
  "c.dose_level_units_original, c.dose_level_units_normalized, ",
  ## conc_medium dictionary fields
  "i.conc_medium_original, i.conc_medium_normalized, ",
  ## administration_route dictionary fields
  "h.administration_route_original, h.administration_route_normalized, ",
  ## administration_method dictionary fields
  # "g.administration_method_original, g.administration_method_normalized, ",
  ## administration_form dictionary fields
  # "f.administration_form_original, f.administration_form_normalized, ",

@@

  # # Join to dictionary tables
  # "LEFT JOIN administration_form_dict f ON c.fk_administration_form_id = f.id ",
  # "LEFT JOIN administration_method_dict g ON c.fk_administration_method_id = g.id ",
  "LEFT JOIN administration_route_dict h ON c.fk_administration_route_id = h.id ",
  "LEFT JOIN conc_medium_dict i ON b.fk_conc_medium_id = i.id ",
  "LEFT JOIN dose_frequency_dict j ON c.fk_dose_frequency_id = j.id ",

  # Rename chemical fields for dosed vs. analyzed chemical record foreign keys
  "LEFT JOIN (SELECT id, dsstox_substance_id AS dosed_chem_dtxsid, ",
  "chemical_name_original AS dosed_chem_name_original, ",
  "chemical_name_secondary_original AS dosed_chem_name_secondary_original, ",
  "dsstox_casrn AS dosed_chem_casrn, preferred_name AS dosed_chem_name ",
  "FROM chemicals) AS k ON c.fk_dosed_chemical_id = k.id ",
  "LEFT JOIN (SELECT id, dsstox_substance_id AS analyzed_chem_dtxsid, ",
  "chemical_name_original AS analyzed_chem_name_original, ",
  "chemical_name_secondary_original AS analyzed_chem_name_secondary_original, ",
  "dsstox_casrn AS analyzed_chem_casrn, preferred_name AS analyzed_chem_name ",
  "FROM chemicals) AS l ON b.fk_analyzed_chemical_id = l.id ",

  # Filtering steps
  "WHERE h.administration_route_normalized IN ('iv', 'oral') AND ",
  "(i.conc_medium_normalized IN ('blood', 'plasma') OR ",
  "i.conc_medium_original IN ('blood', 'plasma')) AND ",
  "(j.dose_frequency_normalized IN ('Once') OR e.curation_set_tag IN ('CVT_Showa'))"
)

@@

# Need to change the column name for this
names(sql_query)[1] <- "conc_time_id"

cvtdb_original <- sql_query
save(cvtdb_original, file = "data/cvtdb_original.rda", compress = "bzip2")
```

```{r harmonizing}
load("data/cvtdb_original.rda")
sql_query <- cvtdb_original
rm(cvtdb_original)

# Harmonize values for easier analysis
cvt_df <- sql_query %>%
  mutate(
    # Note that sometimes fk_test_chemical_id is NA
    across(all_of(c(
      "time_original",
      "time_hr",
      "conc_original",
      "conc_sd_original",
      "conc",
      "conc_sd",
      "dose_level_original",
      "dose_level_normalized",
      "loq"
    )), as.numeric),
    # conc, conc_sd_original, and conc_original
    # have NA/NE/ND values that are coerced to NAs
    year = as.character(year),
    species = tolower(species),
    sex = tolower(sex),
    across(contains("chem_name"), \(x) {
      stringr::str_to_lower(x) |>
        force_ascii_quotes() |>
        # trim whitespace around hyphens
        stringr::str_replace_all("\\s*-\\s*", "-") |>
        # comma-separate forms like 3'6 into 3',6
        stringr::str_replace_all("(?<=\\d)'(?=\\d)", "',") |>
        # fix some bracket cases
        stringr::str_replace_all("\\[3\\]h", "[3h]") |>
        stringr::str_replace_all("\\[a\\]pyr", "(a)pyr") |>
        stringr::str_remove_all(" \\(aids\\)$") |>
        # adds a space after square bracket separator
        stringr::str_replace_all("](?!\\s|-)", "] ")
    }),
    dose_volume_units = ifelse(
      dose_volume_units %in% c("NR", "ND"),
      NA_character_,
      gsub(" / ", "/", gsub("ml", "mL", dose_volume_units))
    ),
    dose_volume = signif(as.numeric(dose_volume), digits = 7),
    dose_duration_units = trimws(dose_duration_units),
    dose_duration = as.numeric(stringr::str_extract(dose_duration,
      pattern = "^\\d+"
    )),
    dose_frequency_normalized = "Once", # Showa is not annotated
    radiolabeled = ifelse(is.na(radiolabeled), FALSE, as.logical(radiolabeled)),
    # for n_subject ranges, take integer value of median
    # otherwise pull number or NA
    n_subjects_normalized = case_when(
      n_subjects_in_series %in% "NR" ~ NA_real_,
      grepl("-", n_subjects_in_series) ~ floor(mean(
        as.numeric(unlist(strsplit(n_subjects_in_series, split = "-"))),
        na.rm = TRUE
      )),
      .default = as.numeric(
        stringr::str_extract(n_subjects_in_series, pattern = "\\d+")
      )
    )
  )
```

## Some data malformations
